### 2021-08-04 chemev/MWbimodality/paper/draft2/age_oh.tex

\documentclass[draft2.tex]{subfiles}
\begin{document}

\section{The Age-\oh~Relation} 
\label{sec:age_oh_relation}

Fig.~\ref{fig:age_oh_comparison} presents a comparison of the age-\oh~ 
relation predicted by our inside-out, late-burst, and outer-burst SFHs to 
the~\citet{Feuillet2019} measurements in the same Galactic regions as in 
Fig.~\ref{fig:amr_insideout_vs_lateburst_fe}. 
The age-\oh~relation shows a smoother population-averaged trend than the 
age-\feh~relation (see Fig.~\ref{fig:amr_insideout_vs_lateburst_fe}). 
Affected by the variability in Type Ia supernova rates discussed in 
\S~\ref{sec:obs_comp:gradient}, the gas-phase Fe abundance at fixed radius 
fluctuates as a function of simulation time, resulting in higher intrinsic 
scatter in the age-\feh~relation than in age-\oh. 
We can make similar arguments about the age-\oh~relation as we do for 
age-\feh~in~\S~\ref{sec:obs_comp:amr}: the late-burst model better reproduces 
the C-shaped nature of the AMR throughout the disc, particularly beyond the 
solar neighbourhood. 
The bin-by-bin comparison is also somewhat more convincing in age-\oh~than in 
age-\feh. 
The late-burst model improves the agreement in all annuli with the exception of 
the solar annulus, where it potentially worsens the agreement slightly, but 
both models adequately reproduce the data there anyway. 
% \par 
In comparing the late-burst model to the outer-burst model, it is clear the 
outer-burst model mitigates the very young ages of the most metal-rich stars 
at~\rgal~= 5 - 7 kpc seen in the late-burst model; this is a consequence of 
the starburst producing a bimodal age distribution at these abundances (see 
discussion in~\S~\ref{sec:obs_comp:amr}). 
% No model illustrated here reproduces the data perfectly, but we have made 
% simple assumptions regarding the enhancement in the SFR. 
% Although these assumptions are nonetheless motivated by the observations of 
% \citet{Mor2019} and~\citet{Isern2019}, the detailed form of the AMR in these 
% models is sensitive to both the strength and time-dependence of the recent SFH 
% as a function of both Galactocentric radius and time. 
% Ascertaining a best-fit set of parameters to describe the burst would require 
% running many variations of our models, perhaps within a more sophisticated 
% mathematical framework.

\end{document}