### 2021-08-04 chemev/MWbimodality/paper/draft2/ms.tex

\documentclass[fleqn, usenatbib]{mnras} 
\usepackage{newtxtext, newtxmath} 	% Required for MNRAS 
\usepackage[T1]{fontenc} 
\usepackage{graphicx}
\usepackage{amsmath}
% \usepackage{amssymb}
\usepackage{mathtools}
\usepackage{multicol}
\usepackage{bm}
\usepackage{pdflscape}
\usepackage{natbib}
\usepackage[section]{placeins}
\usepackage{lipsum}
\usepackage{etoolbox} 
\usepackage{tabularx} 
\usepackage{xcolor} 
\usepackage[toc, page]{appendix} 
\usepackage{subfiles} 
% \voffset-0.6in 
% \usepackage[margin = 0.5in]{geometry} 
\usepackage{geometry} 
\geometry{margin=0.6in} 
% \addtolength{\topmargin}{-0.6in} 
% \hypersetup{draft} 
% \linespread{1.8} 
\hypersetup{ % all blue links 
	colorlinks		= true, 
	urlcolor 		= blue, 
	linkcolor 		= blue, 
	citecolor 		= blue 
}

% recurring expressions 
\newcommand{\ddfrac}[2]{\frac{\displaystyle #1}{\displaystyle #2}} 
\newcommand{\refp}[1]{(\ref{#1})} 
\newcommand{\msun}{\ensuremath{\text{M}_\odot}} 
\newcommand{\persqkpc}{\ensuremath{\text{kpc}^{-2}}} 
\newcommand{\rgal}{\ensuremath{R_\text{gal}}} 
\newcommand{\absz}{\ensuremath{\left|z\right|}} 
\newcommand{\afe}{[$\alpha$/Fe]} 
\newcommand{\ofe}{[O/Fe]} 
\newcommand{\feh}{[Fe/H]} 
\newcommand{\oh}{[O/H]} 
\newcommand{\hsim}{\texttt{h277}} 
\newcommand{\vice}{\texttt{VICE}}

% control over bibtex's alphabetizing 
\providecommand{\noopsort}[1]{}

% figures for the paper in here 
% \graphicspath{ {../plots/} }

\interfootnotelinepenalty = 10000 
\defcitealias{Kroupa2001}{Kroupa} 
\defcitealias{Salpeter1955}{Salpeter}

\title[Stellar Migration and Chemical Enrichment]{Stellar Migration and 
Chemical Enrichment in the Milky Way Disc: A Hybrid Model}

\author[J.W. Johnson et al.]{James W. Johnson,$^{1}$\thanks{
	Contanct e-mail: \href{mailto:
	johnson.7419@osu.edu}{johnson.7419@osu.edu}} 
	David H. Weinberg,$^{1, 2, 3}$ 
	Fiorenzo Vincenzo,$^{1, 2}$ 
	Jonathan C. Bird,$^{4}$ 
	\newauthor 
	Sarah R. Loebman,$^{5}$ 
	Alyson M. Brooks,$^{6}$ 
	Thomas R. Quinn,$^{7}$ 
	Charlotte R. Christensen,$^{8}$ 
	\newauthor 
	and Emily J. Griffith$^{1}$ 
	\\ \null \\ 
	$^{1}$ Department of Astronomy, The Ohio State University, 
	140 W. 18th Ave., Columbus, OH, 43210, USA 
	\\ 
	$^{2}$ Center for Cosmology and Astroparticle Physics (CCAPP), 
	The Ohio State University, 191 W. Woodruff Ave., Columbus, OH, 43210, USA 
	\\ 
	$^{3}$ Institute for Advanced Study, 1 Einstein Dr., Princeton, NJ, 08540, 
	USA 
	\\ 
	$^{4}$ Department of Physics \& Astronomy, Vanderbilt University, 
	2301 Vanderbilt Place, Nashville, TN, 37235, USA 
	\\ 
	$^{5}$ Department of Physics, University of California Merced, 
	5200 North Lake Rd., Merced, CA, 95343, USA 
	\\ 
	$^{6}$ Department of Physics \& Astronomy, Rutgers University, 136 
	Frelinghuysen Rd., Piscataway, NJ, 08854, USA 
	\\ 
	$^{7}$ Department of Astronomy, University of Washington, Box 351580, 
	Seattle, WA, 98195, USA 
	\\ 
	$^{8}$ Department of Physics, Grinnell College, 1116 8th Ave., Grinnell, 
	IA, 50112, USA 
}

\date{Accepted XXX; Received YYY; in original form ZZZ} 
\pubyear{2021}

\begin{document} 
\label{firstpage} 
\pagerange{\pageref{firstpage}--\pageref{lastpage}} 
\maketitle

\subfile{abstract.tex}

\begin{keywords} 
methods: numerical -- galaxies: abundances, evolution, star formation, stellar 
content 
\end{keywords}

\subfile{intro.tex} 
\subfile{methods.tex} 
\subfile{comparison_observations.tex} 
\subfile{conclusions.tex} 
\subfile{acknowledgements.tex} 
\subfile{data_availability.tex}

\begin{appendices} 
\subfile{age_oh.tex} 
\subfile{normalize_sfh.tex} 
\subfile{vice.tex} 
\end{appendices}

\bibliographystyle{mnras} 
\bibliography{ms}

\label{lastpage} 
\end{document}